\chapter{Thesis overview}
\label{sec:TOverview}

This thesis describes the measurement of the neutral current Drell-Yan cross-section in the electron channel measured in the forward region of the ATLAS detector. The data samples are collected throughout 2011 during the collisions at $\sqrt{s} = 7$~TeV at LHC. For the unfolding of the detector-level results, the Monte Carlo (MC) simulation based on the collisions conditions during 2011 is used. The theoretical predictions use various parton distribution functions (PDFs) derived from previous experimental data samples.

All the details of the cross-section measurements are described in the corresponding chapters. The highlights of the analysis include the details of the Monte Carlo generation, the technicalities of the event reconstruction and selection, and the distinct features of the ZeeD analysis software, which allowed to speed-up the analysis process by a significant factor.

The focus of this work is the central-forward (CF) variant of the \Zee\ decay, where "central-forward" means that one of the two final-state electrons is detected in the forward calorimeter, while the other in the central. While the CF analysis share a lot with the central-central (CC) one, which is the other possible variant of the \Zee\ decay (both electrons go to the central parts of the detector), there are still some differences in both the analysis chain and in the results. The combination of the two variants of the analysis is possible, and is done in a paper~\cite{lib:wz2011} that is yet to be published. The work in this thesis contributes to that paper.


\section{Thesis organization}
\label{sec:TOrganization}

The thesis is organized as follows:
\begin{description}
\item Chapter~\ref{sec:Theory}. Brief introduction to the SM, to $pp$-collisions physics and to the $Z$ boson production is given. This chapter also provides a motivation for the \Zee\ cross-section measurements.

\item Chapter~\ref{sec:LHC}. The CERN accelerator complex with LHC is shortly described.

\item Chapter~\ref{sec:ATLAS}. Description of the ATLAS detector and its relevant sub-components is presented. The physics program of the ATLAS experiment is outlined. The ATLAS data acquisition system and computing strategy are discussed as well.

\item Chapter~\ref{sec:DataSamples}. The ATLAS data with its luminosity which was collected in 2011 and used for this thesis is described.

\item Chapter~\ref{sec:MCSamples}. The MC samples generated with different MC algorithms are discussed. The full chain of MC production in the ATLAS experiment is shown. A Frozen Showers system which is the fast simulation system is described, which was developed as part of 2011 MC data preparation.

\item Chapter~\ref{sec:Calibration}. The method of the off-line calibration of the EM calorimeter is explained. The systematic uncertainties in the calibration are discussed as well. Resolution of the EM calorimeters is estimated. Impact of the Fast Simulation techniques is discussed.

\item Chapter~\ref{sec:Reconstruction}. The ATLAS event reconstruction of the EM objects, including trigger performance, electron reconstruction and identification, is described.

\item Chapter~\ref{sec:ZeeD}. The analysis framework developed for this analysis is described. The method to compress data and speed-up the calculations is shown.

\item Chapter~\ref{sec:Selection}. The event selection for $Z \to e^+e^-$ analysis is discussed. The Central-Forward (CF) selection is introduced.

\item Chapter~\ref{sec:ZeeCrossSec}. The method of the cross-section extraction and the binning definition are presented. Several extrapolation techniques are shown.

\item Chapter~\ref{sec:Efficiency}. The electron trigger, identification, reconstruction and isolation efficiencies are described. The details of the forward electron efficiencies are discussed.

\item Chapter~\ref{sec:Bkg}. Studies of the different backgrounds are presented. Several methods for the background estimation are discussed.

\item Chapter~\ref{sec:Results}. Inclusive, single and double differential cross-section measurements in bins of the di-electron rapidity and mass are presented. The background estimation, efficiency determination, acceptance calculation and systematic uncertainties are discussed.

\item Chapter~\ref{sec:Summary}. A summary of the main results is given.
\end{description}
